\huge Tor's Usability for Censorship Circumvention

### 2016-04-26 pets2017-paper/pets2017-paper.tex
serge's suggestions.

We evaluate, redesign, and test the Tor configuration interface by
measuring participants' interactions with the interface in various censorship environments.
The first user study is a small-scale, one-on-one experiment that collected 
behavioral patterns and failure cases with the interface through user observations
and interviews (Section~\ref{sec:qualitative}). This feedback was used to make changes to the interface (Section~\ref{redesign}). The second user study is a large-scale, many-to-one experiment
that collected data on user interactions with the interface to quantify the impact of the design changes (Section~\ref{sec:quantitative}).

@@

There have been been no published usability evaluations of
Tor Browser since the release of the 3.5 series in 2013, which introduced radical UI changes~\cite{torbrowser-35}.
The most recent effort is an unpublished pilot study by Lee and Fifield~\cite{uxsprint} 
that tested the downloading, installation, and browsing tasks in Tor Browser.  This study uncovered a number of issues~\cite{uxsprint2015-tickets},
some of which influenced changes in Tor Browser version~4.5 and later.

@@

In principle, the process of configuring a proxy and bridge can be automated, but
the interface eschews automatic configuration through network probing
in favor of guided manual configuration to give users agency in configuring their connection.
A knowledgeable user can minimize their network trace and hide that they are connecting to Tor.
However, this requires the user to grapple with technical concepts such as 
bridges, pluggable transports, and proxies.

@@

\begin{figure}
\small
\begin{itemize}
\renewcommand{\labelitemi}{$\circ$}
\item Connect with provided bridges\\
Transport type:
\begin{itemize}
\item flashproxy
\item fte
\item fte-ipv6
\item meek-amazon
\item meek-azure
\item meek-google
\item obfs3 (recommended)
\item obfs4
\item scramblesuit
\end{itemize}
\item Enter custom bridges\qquad\fbox{Help}\\
\end{itemize}
\caption{
Bridge selection options in Tor Launcher, as it appeared in Tor Browser~5.0.3.
The options under ``Transport type'' denote various
built-in circumvention technologies (``pluggable transports'').
Under ``Enter custom bridges,'' there is a space to paste in
a bridge line, obtained out of band.
The ``Help'' button displays instructions on obtaining
bridge lines.
}
\label{fig:bridge-options}
\end{figure}

@@

\begin{enumerate}
    \item {\bfseries Task completion rate:} What percentage of users can successfully connect to Tor in a given condition.
    \item {\bfseries  Time to completion:} How long it takes between program startup and a successful Tor bootstrap, which includes time spent searching online for help and waiting for a the connection to bootstrap.
    \item {\bfseries Active configuration time:} How much time a user spends actively configuring network components to connect to Tor, which only includes time spent interacting with the interface, excluding the time waiting for the connection to bootstrap.
\end{enumerate}

@@

\begin{description}
\item[E1: Mild censorship]
The E1 environment only blocks some websites,
as if by DNS poisoning.
We blocked subdomains of torproject.org
and wikipedia.org.
(Reading a page on en.wikipedia.org was the participants' main goal.)
To succeed in this environment,
participants had only to click ``Connect'' on the first screen;
additional configuration was possible but unnecessary.
% (Representative of countries such as France and Australia.)

\item[E2: Intermediate censorship]
The E2 environment blocks websites as in E1;
it additionally blocks the IP addresses of public Tor relays
and directory authorities, simulating a censor that knows
about circumvention and tries to stop it.
In this environment, simply clicking ``Connect'' will not work;
participants had to select a built-in bridge
(any type other than ``flashproxy'' would work).
A custom bridge would work but was not necessary.
% (Representative of countries such as Tunisia.)

\item[E3: Comprehensive censorship]
The E3 environment blocks websites and Tor relays as in E1 and E2;
and additionally blocks the built-in default bridge IP addresses
found in the Tor Browser source code.
To succeed in this environment, participants had either to select
one of the ``meek'' built-in bridges (no other type would work),
or to acquire their own custom bridge and enter it manually.
% (Representative of countries such as China and Syria.)
\end{description}

@@

\subsection{Recruitment}
We recruited 124 participants, about 20 for each
condition.
Data analysis considerations deem 20 participants per combination a 
minimum to assume normal distributions, and 60 participants
per interface revision is adequate to have a meaningful effect size.

We recruited half of our users from Craigslist, and half of our participants from 
the Xlab participant pool. Although Xlab participants are not limited to UC Berkeley students and staff,
a majority of the participants are from campus. For this reason, we chose to recruit 
half of our participants from Craigslist to ensure a diverse set of participants. 
The recruitment text vaguely suggests testing a piece of software, and does not require
that users provide information in advance (Appendix~\ref{quantitative-recruitment}). 
Out of our 124 participants, 59 were recruited from the Xlab pool and the other 65 were
recruited from Craigslist. Ages ranged from 18 to~68 years
($\mu = 28.9$, $\sigma = 12$). 56.8\% were male and 
84.8\% of our participants had at least a college education.

@@

\begin{table}
\centering
% Generated by attempts.R, and manually reordered:
\begin{tabular}{r c c c c c c}
& \rotatebox{90}{E1-NEW} & \rotatebox{90}{E1-OLD} & \rotatebox{90}{E2-NEW} & \rotatebox{90}{E2-OLD} & \rotatebox{90}{E3-NEW} & \rotatebox{90}{E3-OLD} \\
no bridge, no proxy & 17 & 13 &  &  &  &  \\
obfs3, no proxy & 2 & 6 & 18 & 16 &  &  \\
meek-amazon, no proxy &  &  &  &  & 7 & 4 \\
meek-google, no proxy &  &  &  &  & 5 & 4 \\
meek-azure, no proxy &  &  &  &  & 1 & 1 \\
no bridge, 3rd-party proxy &  &  &  &  &  & 1 \\
DNF &  &  & 1 & 3 & 7 & 10 \\
\end{tabular}
\caption{
Bridge--proxy combinations that led to the first successful bootstrap
in each environment and interface.
Most E1 participants used a direct connection,
but a few tried a built-in obfs3 bridge.
All the E2 participants who succeeded,
did so with obfs3 (the recommended bridge type)---none tried
a different bridge before obfs3.
All of the successful E3 participants but one
used one of the meek bridges.
The remaining E3 participant succeeded in an unexpected way:
by searching the web for an open proxy and configuring it
as the proxy setting.
}
\label{tab:attempts-bridge-proxy}
\end{table}

@@

While we did not find a statistically significant difference between the overall time to success between the old and new versions of the interface, the time to success was largely dominated by the time spent waiting to connect to Tor, rather than actively configuring the interface. Table ~\ref{table:overall_time} shows the time spent on each screen while participants used the interface, which shows that participants spent 57\% of their overall time at the progress screen. Participants using the new interface and old interface spent approximately the same amount of time overall on the progress screen (359 minutes versus 339 minutes), which accounts for 78\% and 45\% of their overall time, respectively. We attribute this to a combination of instructions that tell the user to ``wait a couple minutes'' because connections usually take a while to bootstrap and the lack of timeout on the progress screen which keeps the users waiting indefinitely, which were present in both versions of the interface.

The new interface results a statistically significant reduction in the overall active configuration time (Mann--Whitney $Z = -3.055$, $p = 0.00225$, $\mbox{effect size} = 0.284$), while maintaining the same probability for success for the user. We believe that getting rid of the technical bridge and proxy questions, instructing participants that they did not need a proxy, and simplifying the interface attributed to the reduction in time. 
%The impact of the environment on active time is significant (Kruskal--Wallis $\chi^2 = 51.978$, $\mbox{df} = 2$, $p = 10^{-11}$)

@@

\subsubsection{Observations}
These are general observations made from our data set of participants, which may have biases.\\

@@

% impacts of making the proxy page ``automated'' 
\item {\bfseries Users wasted time configuring proxies.} Users believed that they needed proxies to connect. This confusion prevented users from configuring their bridges correctly. Participants with the new interface who were explicitly told that they did not need a proxy spent 31\% of their active time configuring proxies and participants in the old interface who did not receive feedback spent 63\%.

@@

%error messages
\item {\bfseries There were (almost) no error messages.} Many waited for minutes at the progress screen, but never saw an error message. There were warnings, but those warnings went largely unnoticed by our participants. Across both versions of the interface, only one participant ever saw an error message.

% progress bar
\item {\bfseries People waited a long, long time at the progress bar.}
Some dutifully followed the directions to wait, while others are lost for what to do and choose to wait in frustration. Either way, we think that people spent too much time waiting. 
\end{description}

@@

We discarded data on participants who chose to download their own version of Tor Browser online. Some downloaded their own version without trying the provided one on their desktop, while others did so after feeling frustrated with the provided version. Appendix~\ref{tor-downloads} gives the unfortunate details of where Tor was downloaded when torproject.org was blocked.

@@

\begin{description}
\item{\bfseries Automate the configuration.} Complete automation of the configuration process, which involves network probing, would not harm users today (as far as we know). Our study finds that most users fail on the first try when trying to circumvent an advanced censor, so this automation may not increase the number of users who leak that they are connecting to the Tor network. 
\item{\bfseries Ask about the risk.} Automating the configuration process for all who are not at risk and offering manual configuration to those who would like to be more cautious is an alternative to complete automation. The complication with this approach is that we would need to ask the users about the risk involved with connecting to the Tor network. Users may not be qualified to answer this question or may not trust Tor with this information. 
\item{\bfseries Ask if users if they know what to do.} We ask users if they know how to configure a bridge and proxy, and automate the configuration process if they do not. However, there are ethical implications of making mistakes on behalf of the users if there is risk involved. 
\end{description}

@@

We test the impact of our changes on 116 participants in a quantitative user study to find that that the changes cause a statistically significant reduction in the time participants spent configuring their connection, with no decrease in success rates. We observe and quantify common behavioral trends based on logs and video recordings of our participants, and offer recommendations that we believe will greatly increase the success rate and time to success for future users.

>>>>

We evaluate, redesign, and test the Tor configuration interface by
measuring participants' interactions with the interface in various censorship environments.
The first user study is a small-scale, qualitative experiment that collected 
behavioral patterns and failure cases with the interface through user observations
and interviews (Section~\ref{sec:qualitative}). This feedback was used to make changes to the interface (Section~\ref{redesign}). The second user study is a large-scale, quantitative experiment
that collected data on user interactions with the interface to quantify the impact of the design changes (Section~\ref{sec:quantitative}).

@@

We do not know of any published usability evaluations of
Tor Browser since the release of the 3.5 series in 2013, which introduced radical UI changes~\cite{torbrowser-35}.
The most recent effort is an unpublished pilot study by Lee and Fifield~\cite{uxsprint} 
that tested the downloading, installation, and browsing tasks in Tor Browser.  This study uncovered a number of issues~\cite{uxsprint2015-tickets},
some of which influenced changes in Tor Browser version~4.5 and later.

@@

In principle, the process of configuring a proxy and bridge can be automated, but
the interface eschews automatic configuration through network probing
in favor of guided manual configuration to give users agency in configuring their connection. 
Automatic configuration through network probing may put some users in certain regimes at risk. 
A knowledgeable user can minimize their network trace and hide that they are connecting to Tor.
However, this requires the user to grapple with technical concepts such as 
bridges, pluggable transports, and proxies.

@@

\begin{figure}
\small
\begin{itemize}
\renewcommand{\labelitemi}{$\circ$}
\item Connect with provided bridges\\
Transport type:
\begin{itemize}
\item flashproxy
\item fte
\item fte-ipv6
\item meek-amazon
\item meek-azure
\item meek-google
\item obfs3 (recommended)
\item obfs4
\item scramblesuit
\end{itemize}
\item Enter custom bridges\qquad\fbox{Help}\\
\end{itemize}
\caption{
Bridge selection options in Tor Launcher, as it appeared in Tor Browser~5.0.3.
Users are not familiar with the options under ``Transport type'' denote various
built-in circumvention technologies (``pluggable transports'').
Under ``Enter custom bridges,'' there is a space to paste in
a bridge line, obtained out of band.
The ``Help'' button displays instructions on obtaining
bridge lines. 
}
\label{fig:bridge-options}
\end{figure}

@@

\begin{enumerate}
    \item {\bfseries Task completion rate:} What percentage of users can successfully connect to Tor in a given condition.
    \item {\bfseries  Time to completion:} How long it takes between program startup and a successful Tor bootstrap, which includes time spent searching online for help and waiting for the connection to bootstrap.
    \item {\bfseries Active configuration time:} How much time a user spends actively configuring network components to connect to Tor, which only includes time spent interacting with the interface, excluding the time waiting for the connection to bootstrap.
\end{enumerate}

@@

% E1
% (Representative of countries such as France and Australia.)
% E2 
% (Representative of countries such as Tunisia.)
% E3
% (Representative of countries such as China and Syria.)

\begin{description}
\item[E1: Mild censorship]
(Representative of countries such as France and Australia.)
The E1 environment only blocks some websites,
as if by DNS poisoning.
We blocked subdomains of torproject.org
and wikipedia.org.
(Reading a page on en.wikipedia.org was the participants' main goal.)
To succeed in this environment,
participants had only to click ``Connect'' on the first screen;
additional configuration was possible but unnecessary.
% (Representative of countries such as France and Australia.)

\item[E2: Intermediate censorship]
(Representative of countries such as Tunisia.)
The E2 environment blocks websites as in E1;
it additionally blocks the IP addresses of public Tor relays
and directory authorities, simulating a censor that knows
about circumvention and tries to stop it.
In this environment, simply clicking ``Connect'' will not work;
participants had to select a built-in bridge
(any type other than ``flashproxy'' would work).
A custom bridge would work but was not necessary.
% (Representative of countries such as Tunisia.)

\item[E3: Comprehensive censorship]
(Representative of countries such as China and Syria.)
The E3 environment blocks websites and Tor relays as in E1 and E2;
and additionally blocks the built-in default bridge IP addresses
found in the Tor Browser source code.
To succeed in this environment, participants had either to select
one of the ``meek'' built-in bridges (no other type would work),
or to acquire their own custom bridge and enter it manually.
% (Representative of countries such as China and Syria.)
\end{description}

@@

\subsection{Recruitment}
We recruited 124 participants, about 20 for each
condition. We recruited half of our users from Craigslist, and half of our participants from 
the Xlab participant pool. Although Xlab participants are not limited to UC Berkeley students and staff,
a majority of the participants are from campus. For this reason, we chose to recruit 
half of our participants from Craigslist to ensure a diverse set of participants. 
The recruitment text vaguely suggests testing a piece of software, and does not require
that users provide information in advance (Appendix~\ref{quantitative-recruitment}). 
Out of our 124 participants, 59 were recruited from the Xlab pool and the other 65 were
recruited from Craigslist. Ages ranged from 18 to~68 years
($\mu = 28.9$, $\sigma = 12$). 56.8\% were male and 
84.8\% of our participants had at least a college education.

@@

The time to success was largely dominated by the time spent waiting to connect to Tor, rather than actively configuring the interface. Table ~\ref{table:overall_time} shows the time spent on each screen while participants used the interface, which shows that participants spent 57\% of their overall time at the progress screen. Participants using the new interface and old interface spent approximately the same amount of time overall on the progress screen (359 minutes versus 339 minutes), which accounts for 78\% and 45\% of their overall time, respectively. We attribute this to a combination of instructions that tell the user to ``wait a couple minutes'' because connections usually take a while to bootstrap and the lack of timeout on the progress screen, which were present in both versions of the interface.

The new interface results a statistically significant reduction in the overall active configuration time (Mann--Whitney $Z = -3.055$, $p = 0.00225$, $r= 0.284$), while maintaining the same probability for success for the user. We believe that getting rid of the technical bridge and proxy questions, instructing participants that they did not need a proxy, and simplifying the interface attributed to the reduction in time. 
%The impact of the environment on active time is significant (Kruskal--Wallis $\chi^2 = 51.978$, $\mbox{df} = 2$, $p = 10^{-11}$)

@@

\begin{table}
\centering
% Generated by attempts.R, and manually reordered:
\begin{tabular}{r c c c c c c}
& \rotatebox{90}{E1-NEW} & \rotatebox{90}{E1-OLD} & \rotatebox{90}{E2-NEW} & \rotatebox{90}{E2-OLD} & \rotatebox{90}{E3-NEW} & \rotatebox{90}{E3-OLD} \\
no bridge, no proxy & 17 & 13 &  &  &  &  \\
obfs3, no proxy & 2 & 6 & 18 & 16 &  &  \\
meek-amazon, no proxy &  &  &  &  & 7 & 4 \\
meek-google, no proxy &  &  &  &  & 5 & 4 \\
meek-azure, no proxy &  &  &  &  & 1 & 1 \\
no bridge, 3rd-party proxy &  &  &  &  &  & 1 \\
DNF (did not finish) &  &  & 1 & 3 & 7 & 10 \\
\end{tabular}
\caption{
Bridge--proxy combinations that led to the first successful bootstrap
in each environment and interface.
Most E1 participants used a direct connection,
but a few tried a built-in obfs3 bridge.
All the E2 participants who succeeded,
did so with obfs3 (the recommended bridge type)---none tried
a different bridge before obfs3.
All of the successful E3 participants but one
used one of the meek bridges.
The remaining E3 participant succeeded in an unexpected way:
by searching the web for an open proxy and configuring it
as the proxy setting.
}
\label{tab:attempts-bridge-proxy}
\end{table}

We believe that our change to add advice on what to try next on error messages (such as try the connection again, reconfigure your bridge settings, or try without a proxy) would have improved the rate of success. However, only 1 of 116 participants saw an error during the experiment. That participant, who used the new interface, saw that the bridge configuration was the source of the error, and heeded the interface's advice to reconfigure the bridge options. But since most participants did not see any error messages, any effects they may have had on success rates are not reflected in our results. The addition of a summary screen did cause behavioral changes (13\% of participants saw their current settings, then went back to adjusted them), but this behavioral change did not have a statistically significant impact on success rates. We added pre-preemptive advice to the bridge configuration screen to first try an obfs3 bridge and then a meek bridge, but we suspect that most participants did not read this advice.

@@

\subsubsection{Observations}
These are general observations made from our data set of participants, which may or may not be representative of the Tor user base.\\

@@

% impacts of making the proxy page ``automated'' 
\item {\bfseries Users wasted time configuring proxies.} Users believed that they needed proxies to connect. This confusion prevented users from configuring their bridges correctly. Participants with the new interface who were explicitly told that they did not need a proxy spent 31\% of their active time configuring proxies, perhaps due to not reading the text or not believing the interface. Participants in the old interface who did not receive feedback spent 63\%.

@@

%error messages
\item {\bfseries There were (almost) no error messages.} Many waited for minutes at the progress screen, but never saw an error message. Error messages are intended to appear after a timeout. There were warnings, but those warnings went largely unnoticed by our participants. Across both versions of the interface, only one participant ever saw an error message.

% progress bar
\item {\bfseries People waited a long, long time at the progress bar.}
From watching the screen capture videos, we see that some participants dutifully followed the directions to wait, while others are lost for what to do and choose to wait in frustration.
\end{description}

@@

We discarded data on participants who chose to download their own version of Tor Browser online. 5 participants downloaded their own version without trying the provided one on their desktop, while others did so after feeling frustrated with the provided version. Appendix~\ref{tor-downloads} gives the unfortunate details of where Tor was downloaded when torproject.org was blocked.

@@

\begin{description}
\item{\bfseries Automate the configuration.} Complete automation of the configuration process, which involves network probing, would not harm users today (as far as we know). Our study finds that most users fail on the first try when trying to circumvent an advanced censor, so this automation may not increase the number of users who leak that they are connecting to the Tor network. 
\item{\bfseries Ask about the risk.} Automating the configuration process for all who are not at risk and offering manual configuration to those who would like to be more cautious is an alternative to complete automation. The complication with this approach is that we would need to ask the users about the risk involved with connecting to the Tor network. Users may not be qualified to answer this question or may not trust Tor with this information. 
\item{\bfseries Ask if users know what to do.} Automating the process for users who do not know how to configure a bridge or proxy can be an alternative option. However, there are ethical implications of making mistakes on behalf of users with risk involved. 
\end{description}

@@

We test the impact of our changes on 116 participants in a quantitative user study to find that the changes cause a statistically significant reduction in the time participants spent configuring their connection, with no decrease in success rates. We observe and quantify common behavioral trends based on logs and video recordings of our participants, and offer recommendations that we believe will greatly increase the success rate and time to success for future users.
